Species association networks unveil scale effects in community assemblage mechanisms.

### 2025-07-23 PAPER_CODE.qmd
removed deprecated and change anablepsoides to rivulus

```{r}
log_abundance <- log(1 + NR_DATA$Abundance)
log_abundance_sorted <- log_abundance[order(NR_DATA$stream), ]

@@

pheatmap(log_abundance_sorted,
         color = color_palette,
         cluster_rows = FALSE, cluster_cols = FALSE,
         annotation_row = data.frame(Stream = NR_DATA$stream),
         show_rownames = FALSE)
```
## Covariates preparation (PCA)

@@

# Cross-validation measures for PLN-network and ZIPLN-network
```{r cross-val-folds}
K = 5
set.seed(1)
folds    = split(sample(1:nrow(covar_data)), rep(1:K, length = nrow(covar_data)))
```

@@

## PLN-network cross-validation
```{r}
cv_pln_out      <- crossval(Abundance ~ 1 + PC1 + PC2 + stream + disturbance,
                         data = NR_DATA)
print(median(cv_pln_out$err))

@@

## ZIPLN-network cross-validation
```{r}
cv_pln_out_ZI   <- crossval(Abundance ~ 1 + PC1 + PC2 + disturbance | stream ,
                            data = NR_DATA, method = "ZIPLN")
print(median(cv_pln_out_ZI$err))

@@

covariates = cbind(NR_COVARIATES, NR_PCA_SCORES)
NR_DATA <- prepare_data(
  counts     = NR_COUNTS,
  covariates = covariates
)
covar_data <- NR_DATA %>% select(c("PC1", "PC2", "disturbance", "stream"))
```
## BIC

>>>>

```{r}
log_abundance <- log(1 + NR_COUNTS)
log_abundance_sorted <- log_abundance[order(NR_COVARIATES$stream), ]

@@

pheatmap(log_abundance_sorted,
         color = color_palette,
         cluster_rows = FALSE, cluster_cols = FALSE,
         annotation_row = data.frame(Stream = NR_COVARIATES$stream),
         show_rownames = FALSE)
```

## Covariates preparation (PCA)

@@

# Cross-validation measures for PLN-network and ZIPLN-network

```{r cross-val-folds}
K = 5
set.seed(1)
folds    = split(sample(1:nrow(NR_DATA)), rep(1:K, length = nrow(NR_DATA)))
```

@@

## PLN-network cross-validation

```{r}
cv_pln_out      <- crossval(Abundance ~ 1 + PC1 + PC2 + stream + disturbance,
                         data = NR_DATA)
print(median(cv_pln_out$err))

@@

## ZIPLN-network cross-validation

```{r}
cv_pln_out_ZI   <- crossval(Abundance ~ 1 + PC1 + PC2 + disturbance | stream ,
                            data = NR_DATA, method = "ZIPLN")
print(median(cv_pln_out_ZI$err))

@@

covariates = cbind(NR_COVARIATES, NR_PCA_SCORES)
NR_DATA <- prepare_data(
  counts     = NR_COUNTS,
  covariates = covariates
)
covar_data <- NR_DATA %>% select(c("PC1", "PC2", "disturbance", "stream"))
```

## BIC
